Self recovering capabilities in images using the BPCS Steganography algorithm

### 2014-10-09 Paper/main.tex
Section 4 changes and diagrams added along with other minor changes

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[T1]{fontenc}
\usepackage[margin=1.0in]{geometry}
\usepackage{mathtools}
\usepackage{setspace}
\setstretch{1.0}

@@

\noindent Abstract of this paper and what I wrote in it.
This will give the reader a short overview of what they can expect to find in this paper.

@@

\subsection{Introduction}
This paper proposes a novel method for self-recovering images.

@@

\subsection{Terminology}
This paper makes use of several terms and acronyms.
This section serves to make the meaning of those terms clear.

\begin{itemize}
  \item Original Image: This refers to the original image without any modifications.
  \item Embedding: Embedding in the context of this paper refers to the process of storing digital information in an image's content in such a way that the original image is still distinguishable.
  \item Image after embedding: This refers to the the original image after the embedding process has been applied.
  \item Image after reconstruction: This refers to the reconstructed version of the original image that was obtained by applying the reconstruction process to the image after embedding.
  \item LSB: This refers to the least significant bit in a binary string.
  \item MSB: This refers to the most significant bit/bits in a binary string.
  \item Section, Image section, Pixel section: This refers to a section of 8x8 pixels in an image.
  \item Bit plane: A bit plane is a 8x8 grid representation of a specific bit value's for a specific pixel section. Each pixel has 24 bits which in total represents each colour, 8 bits per colour. these bits are the 8 binary bits that represent the numerical value of the colour. These 8 bits are separated from each other and are each stored in a 8x8 grid in the same location as that of the pixel in the pixel section. This process is repeated for each pixel in the pixel section and the result is 8, 8x8 grids of binary numbers. Each one of these 8 grids represent a single bit plane. 
  \item Block: This refers to an individual bit plane of a specific section of the image. There are 8 such blocks in each section since each pixel is represented by 8 bits for each colour.
  \item BPCS Steganography: Bit-Plane Complexity Segmentation Steganography
\end{itemize}

\subsection{Conclusion}
This chapter discussed the purpose behind self-recovering images as well as the aim of this research and the proposed method. 
The research methodology followed was described. A list of terminologies used in this paper was also given.
The next chapter will give a more detailed overview of all the components that make up self-recovering images.

@@

\hspace{0pt} \\
The BPCS algorithm is a high capacity embedding algorithm. The goal of the BPCS algorithm is thus to embed as much data as possible into a cover image without detection \cite{beaullieubpcs}.
This refers to detection by human perception as well as statistical analysis although the latter is not the focus of this paper.
The BPCS algorithm uses an adaptive scheme to classify blocks that are suitable for embedding (similar to \ref{DynamicBlockOverview}).
BPCS steg differs however in the fact that the blocks can be conjugated in a checker board pattern to increase complexity and decrease detectability if needs be.
BPCS steg also differs in the way in which embedding takes place.
Instead of always embedding into the 3-LSB planes as in \ref{DynamicBlockOverview}.
Bit planes are dynamically allocated for embedding.
One section may thus only embed into one bit plane while another may use all 7 available bit planes.

@@

\hspace{0pt} \\
The proposed method in this paper attempts to apply the algorithm for BPCS Steganography
to the problem of self recovering images, specifically to increase overall quality.
The BPCS algorithm can achieve a very high capacity while still remaining undetectable through human perception.
This makes the BPCS algorithm a good choice to increase the capacity of a self recovering image scheme and possibly increase the quality through this process.

@@

\subsection{Overview}
\begin{itemize}
	\item Erasure channel: An erasure channel is a channel that can transmit a bit of information and the receiver can either receive the bit or the bit or it receives a message that the bit was not received, ie the bit was erased.
	\item Peak signal to noise ratio (PSNR): PSNR is an expression for the ratio between the maximum possible value (power) of a signal and the power of distorting noise that affects the quality of its representation. PSNR is represented as a logarithmic decibel scale. In terms of images PSNR can objectively measure the amount of noised introduced into an image when compared with the original image. 
	As the mean squared error between the two images approaches 0 the PSNR approaches infinity. A higher PSNR value thus provides a higher quality image. 
	A small PSNR value implies a high numerical difference between the two images \cite{hore2010image}.  
	The quality measures mentioned in the 3 methods discussed were measured using the PSNR method. \cite{huynh2008scope} However determined that the validity of PSNR is limited to modifications within the same image.
	It is thus not reliable to compare PSNR values between two different images.
	\item Tampering rate: tampering rate is the percentage of image content that can be modified in the image after embedding, while still being able to reconstruct all the modified parts of that image.
	If too much of the image is modified the embedded image content would be lost and reconstruction would not be possible.
	\item Security: security refers to how easy it is to detect that information has been embedded into the image using either visual or statistical methods. Good security implies that it is difficult to detect and poor security implies that it is simple t detect.
	\item Payload size: Payload size is defined as the amount of bits that is embedded in to an image using steganographic embedding techniques. 
	\item Embedding capacity: embedding capacity is defined as the maximum amount of bits that can be embedded into an image using steganographic embedding techniques.
\end{itemize}
//TODO <All the terms that have not been defined yet in the paper that will be used in this section should be defined here, query if this is indeed a correct assumption>

@@

\hspace{0pt} \\
All three methods thus provide good results and would serve as a good benchmark for comparing the method proposed in this paper.

\subsection{Conclusion}
This chapter looked at currently available methods for self-recovering images.
An erasure channel model utilizing the remaining authentic content in section \ref{ErasureChannelOverview}, a method using difference expansion in section \ref{differenceExpansionOverview} and a method using dynamic block allocation \ref{DynamicBlockOverview}.
A comparison between these methods was also provided in section \ref{litStudyCamparison}.
The next chapter will discuss the implementation of the of the self-recovering image method proposed in this paper.

\section{Implementation}
\label{Implementation}
\subsection{Introduction}
//TODO

\subsection{Embedding}
\label{embedding}

The proposed method works with colour images which means that for each pixel there is a red, green and blue value. This means that each block effectively stores 3 bits for each pixel.

@@

\hspace{0pt} \\
As a solution to this problem each bit plane block used for embedding is treated as a separate embedding entity.
There is thus no overlap between two different blocks used for embedding.
The image reference content is first processed using the method explained later in this section.
Each block has 12 pixels (with a bit depth of 16 bits) of the image reference content embedded into it.
An embedding map containing the location of the first pixel to be embedded in the block is also created.
Since the individual image reference content pixels are embedded in order it is only necessary to know the location of the first pixel.
This embedding map is then embedded into the LSB (Least significant bit) layer of that specific 8x8 block of pixels 
This allows the extractor to be able to identify where all the extracted pixels belong in the image even if the image has been modified.
This serves to give random access to the embedded content which is an essential property of self embedding as defined by (Korus et al. \cite{korus2013efficient}).
This is essential in order for the reconstruction process to function properly.

@@

\paragraph{Image resizing}
\label{ImageCompression}
The image reference content that is embedded into the original image cannot be embedded with perfect quality due to space restrictions caused by the embedding algorithm. 
The proposed method however does not utilize compression as it is beyond the scope of this paper.
The proposed method rather reduces the bit rate of the image from 24 bits to 16 which reduces the size of each pixel by 6 bits.
The image's size is also reduced according to the space available for embedding.
When extracting the image reference content again the assembled image reference content is scaled up to it's original size of 512x512 pixels before completing the reconstruction.
This is done to ensure that the pixels contained in the image reference content matches up (location wise) with the pixels in the original image.

@@

\hspace{0pt} \\
The block complexity is calculated by using a border complexity method. \cite{beaullieubpcs}
That is for each binary bit that has an adjacent binary bit that is different than itself the border count is increased. 
The final border count is then divided with the theoretical maximum border count, which would be the border count for a grid of the same size with a checkerboard pattern.

@@

\hspace{0pt} \\
This method of conjugation is a simple method that makes a block more complex while still allowing the original values of the block to be retrieved.

@@

\subsection{Extraction and Reconstruction}
\label{extractRestore}
\subsubsection{Image authentication}
The first step for the image extraction and reconstruction process is to determine which, if any, parts of the image after embedding has been tampered with.
This is accomplished by first completing steps 1-4 of the BPCS steg algorithm (\ref{bpcsSteg}).
for each section (section refers here to 8x8 pixel section, thus a section here contains all 8 bit planes) the authentication hash string is calculated as described in \ref{impFragileWatermark}. 
The embedded hash is then extracted from the LSB plane of that block.
The extracted hash is then compared with the calculated hash and if they match the section is authentic.
If they do not match then that section has been tampered with.

\subsubsection{Image reconstruction}
\hspace{0pt} \\
For each authentic block the 7 MSB planes have their complexity calculated.
If a bit plane block is found to be complex enough it is assumed that there is image information embedded into it.
The embedded conjugation map is extracted at this point and if it is found that any of the complex enough bit plane blocks were conjugated the conjugation algorithm is applied again to reverse the conjugation.

@@

\section{Results} 
\subsection{Introduction}
//TODO

In testing the proposed method a simple experiment was set up in which the original image was processed using the embedding process described in \ref{embedding}.
The results of that process can be seen in figure \ref{fig:lenaEmbedOnceEmbedding}, showing the original image and the image after the embedding process.
Note that in this case the Image Reference content was only embedded once

@@

As an experimental control the results from the experiment in the proposed method will be compared to other current methods for images with self recovering capabilities.
This is discussed further in  section \ref{interpretation}.
\subsection{Conclusion}
//TODO

\section{Interpretation of results}
\subsection{Introduction}
//TODO

>>>>

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[T1]{fontenc}
\usepackage[margin=1.0in]{geometry}
\usepackage{mathtools}
\usepackage{setspace}
\setstretch{1.5}

@@

\noindent Abstract of this dissertation and what I wrote in it.
This will give the reader a short overview of what they can expect to find in this dissertation.

@@

\subsection{Introduction}
This dissertation proposes a novel method for self-recovering images.

@@

\subsection{Terminology}
This dissertation makes use of several terms and acronyms.
This section serves to make the meaning of those terms clear.

\begin{itemize}
  \item Original Image: This refers to the original image without any modifications.
  \item Embedding: Embedding in the context of this dissertation refers to the process of storing digital information in an image's content in such a way that the original image is still distinguishable.
  \item Image after embedding: This refers to the the original image after the embedding process has been applied.
  \item Image after reconstruction: This refers to the reconstructed version of the original image that was obtained by applying the reconstruction process to the image after embedding.
  \item LSB: This refers to the least significant bit in a binary string.
  \item MSB: This refers to the most significant bit/bits in a binary string.
  \item Section, Image section, Pixel section: This refers to a section of 8x8 pixels in an image.
  \item Bit plane: A bit plane is a 8x8 grid representation of a specific bit value's for a specific pixel section. Each pixel has 24 bits which in total represents each colour, 8 bits per colour. these bits are the 8 binary bits that represent the numerical value of the colour. These 8 bits are separated from each other and are each stored in a 8x8 grid in the same location as that of the pixel in the pixel section. This process is repeated for each pixel in the pixel section and the result is 8, 8x8 grids of binary numbers. Each one of these 8 grids represent a single bit plane. 
  \item Block: This refers to an individual bit plane of a specific section of the image. There are 8 such blocks in each section since each pixel is represented by 8 bits for each colour.
  \item BPCS Steganography: Bit-Plane Complexity Segmentation Steganography
\end{itemize}

\subsection{Conclusion}
This chapter discussed the purpose behind self-recovering images as well as the aim of this research and the proposed method. 
The research methodology followed was described. A list of terminologies used in this dissertation was also given.
The next chapter will give a more detailed overview of all the components that make up self-recovering images.

@@

\hspace{0pt} \\
The BPCS algorithm is a high capacity embedding algorithm. The goal of the BPCS algorithm is thus to embed as much data as possible into a cover image without detection \cite{beaullieubpcs}.
This refers to detection by human perception as well as statistical analysis although the latter is not the focus of this dissertation.
The BPCS algorithm uses an adaptive scheme to classify blocks that are suitable for embedding (similar to \ref{DynamicBlockOverview}).
BPCS steg differs however in the fact that the blocks can be conjugated in a checker board pattern to increase complexity and decrease detectability if needs be.
BPCS steg also differs in the way in which embedding takes place.
Instead of always embedding into the 3-LSB planes as in \ref{DynamicBlockOverview}.
Bit planes are dynamically allocated for embedding.
One section may thus only embed into one bit plane while another may use all 7 available bit planes.

@@

\hspace{0pt} \\
The proposed method in this dissertation attempts to apply the algorithm for BPCS steganography
to the problem of self recovering images, specifically to increase overall quality.
The BPCS algorithm can achieve a very high capacity while still remaining undetectable through human perception.
This makes the BPCS algorithm a good choice to increase the capacity of a self recovering image scheme and possibly increase the quality through this process.

@@

\subsection{Overview}
\begin{itemize}
	\item Erasure channel: An erasure channel is a channel that can transmit a bit of information and the receiver can either receive the bit or the bit or it receives a message that the bit was not received, ie the bit was erased.
	\item Peak signal to noise ratio (PSNR): PSNR is an expression for the ratio between the maximum possible value (power) of a signal and the power of distorting noise that affects the quality of its representation. PSNR is represented as a logarithmic decibel scale. In terms of images PSNR can objectively measure the amount of noised introduced into an image when compared with the original image. 
	As the mean squared error between the two images approaches 0 the PSNR approaches infinity. A higher PSNR value thus provides a higher quality image. 
	A small PSNR value implies a high numerical difference between the two images \cite{hore2010image}.  
	The quality measures mentioned in the 3 methods discussed were measured using the PSNR method. \cite{huynh2008scope} However determined that the validity of PSNR is limited to modifications within the same image.
	It is thus not reliable to compare PSNR values between two different images.
	\item Tampering rate: tampering rate is the percentage of image content that can be modified in the image after embedding, while still being able to reconstruct all the modified parts of that image.
	If too much of the image is modified the embedded image content would be lost and reconstruction would not be possible.
	\item Security: security refers to how easy it is to detect that information has been embedded into the image using either visual or statistical methods. Good security implies that it is difficult to detect and poor security implies that it is simple t detect.
	\item Payload size: Payload size is defined as the amount of bits that is embedded in to an image using steganographic embedding techniques. 
	\item Embedding capacity: embedding capacity is defined as the maximum amount of bits that can be embedded into an image using steganographic embedding techniques.
\end{itemize}
//TODO <All the terms that have not been defined yet in the dissertation that will be used in this section should be defined here, query if this is indeed a correct assumption>

@@

\hspace{0pt} \\
All three methods thus provide good results and would serve as a good benchmark for comparing the method proposed in this dissertation.

\subsection{Conclusion}
This chapter looked at currently available methods for self-recovering images.
An erasure channel model utilizing the remaining authentic content in section \ref{ErasureChannelOverview}, a method using difference expansion in section \ref{differenceExpansionOverview} and a method using dynamic block allocation \ref{DynamicBlockOverview}.
A comparison between these methods was also provided in section \ref{litStudyCamparison}.
This information will be used to determine the success of the self-recovering image method proposed in this dissertation by comparing the results from the proposed method with the results achieved by the methods discussed in this chapter.
The next chapter will discuss the implementation of the of the self-recovering image method proposed in this dissertation.

\section{Overview of the proposed system}
\label{Implementation}
\subsection{Introduction}
This chapter discusses the details of the implementation for the self-recovering image method proposed in this dissertation. Embedding is discussed under which BPCS staganography and fragile watermarking will also be discussed.
The extraction and restoration process will be discussed under which authentication and restoration will also be discussed.
Diagrams are also provided to visualize the embedding as well as the extraction process

\subsection{Overview of proposed system}
\label{impOverview}
The self-recovering method proposed in this dissertation uses BPCS steganography to increase the payload capacity of the cover image in order to be able to embed more reference content into the image without introducing visual distortion.
In addition a fragile watermark was used to detect which parts of the image was damaged and thus needs to be reconstructed.

\subsection{Embedding}
\label{embedding}
In the embedding phase BPCS steganography and fragile watermarks were used.  How these two technologies were implemented are discussed in sections \ref{bpcsSteg} and \ref{impFragileWatermark}

\begin{figure}[h]
\centerline{%
\includegraphics[scale=0.6]{"EmbeddingProcess"}%
} %
\caption{The embedding process of the proposed self-recovering method}
\label{fig:embeddingProcess}
\end{figure}

@@

\hspace{0pt} \\
As a solution to this problem each bit plane block used for embedding is treated as a separate embedding entity.
There is thus no overlap between two different blocks used for embedding.
The image reference content is first processed using the method explained later in this section.
Each block has 12 pixels (with a bit depth of 16 bits) of the image reference content embedded into it.
An embedding map containing the location of the first pixel to be embedded in the block is also created.
Since the individual image reference content pixels are embedded in order it is only necessary to know the location of the first pixel.
This embedding map is then embedded into the LSB (Least significant bit) layer of that specific 8x8 block of pixels 
This allows the extractor to be able to identify where all the extracted pixels belong in the image even if the image has been modified.
This serves to give random access to the embedded content which is an essential property of self embedding as defined by Korus et al. \cite{korus2013efficient}.
This is essential in order for the reconstruction process to function properly.

\hspace{0pt} \\
The proposed method works with colour images which means that for each pixel there is a red, green and blue value. This means that each block effectively stores 3 bits for each pixel.

@@

\paragraph{Image resizing}
\label{ImageCompression}
The image reference content that is embedded into the original image cannot be embedded with perfect quality due to space restrictions caused by the embedding algorithm. 
The proposed method however does not utilize compression as it is beyond the scope of this dissertation.
The proposed method rather reduces the bit rate of the image from 24 bits to 16 which reduces the size of each pixel by 6 bits.
The image's size is also reduced according to the space available for embedding.
When extracting the image reference content again the assembled image reference content is scaled up to it's original size of 512x512 pixels before completing the reconstruction.
This is done to ensure that the pixels contained in the image reference content matches up (location wise) with the pixels in the original image.

@@

\hspace{0pt} \\
The block complexity is calculated by using a border complexity method. \cite{beaullieubpcs}.
That is for each binary bit that has an adjacent binary bit that is different than itself the border count is increased. 
The final border count is then divided with the theoretical maximum border count, which would be the border count for a grid of the same size with a checkerboard pattern.

@@

\hspace{0pt} \\
This method of conjugation is a simple method that makes a block more complex while still allowing the original values of the block to be retrieved.
A conjugation map is also created to indicate which blocks have been conjugated.
This is needed since the receiver of the image would not be able to determine if a block has been conjugated or no without such a map.

@@

\subsection{Extraction and Reconstruction}
\label{extractRestore}
In the extraction and reconstruction process of the proposed method the image is first authenticated, this process is discussed in section \ref{extractionAuthentication}. 
The image reference content is then extracted and the image is restored to an approximation of the original image before embedding, this process is discussed in section \ref{extractionReconstruction}.

\begin{figure}[h]
\centerline{%
\includegraphics[scale=0.6]{"ExtractionProcess"}%
} %
\caption{The extraction process of the proposed self-recovering method}
\label{fig:extractionProcess}
\end{figure}

\subsubsection{Image authentication}
\label{extractionAuthentication}
The first step for the image extraction and reconstruction process is to determine which, if any, parts of the image after embedding has been tampered with.
This is accomplished by first completing steps 1-4 of the BPCS steg algorithm (\ref{bpcsSteg}).
for each section (section refers here to 8x8 pixel section, thus a section here contains all 8 bit planes) the authentication hash string is calculated as described in \ref{impFragileWatermark}. 
The embedded hash is then extracted from the LSB plane of that block.
The extracted hash is then compared with the calculated hash and if they match the section is authentic.
If they do not match then that section has been tampered with.

\subsubsection{Image reconstruction}
\label{extractionReconstruction}
For each authentic block the 7 MSB planes have their complexity calculated.
If a bit plane block is found to be complex enough it is assumed that there is image information embedded into it.
The embedded conjugation map is extracted at this point and if it is found that any of the complex enough bit plane blocks were conjugated the conjugation algorithm is applied again to reverse the conjugation.

@@

\subsection{Conclusion}
This chapter discussed the proposed self-recovering system. An overview was given of the proposed method in section \ref{impOverview}. 
Embedding was discussed in section \ref{embedding} under which BPCS steganography as well as fragile watermarking was also discussed in sections \ref{bpcsSteg} and \ref{impFragileWatermark}.
The extraction process was discussed in section \ref{extractRestore} under which authentication and reconstruction was also discussed in sections \ref{extractionAuthentication} and \ref{extractionReconstruction}. 
Diagrams were also provided to visualize the embedding process in figure \ref{fig:embeddingProcess} and the extraction process in figure \ref{fig:extractionProcess}.
The next chapter will discuss the experimental results achieved using the proposed method.

\section{Experimental Results} 
\subsection{Introduction}
//TODO

\subsection{Experiments conducted}
In testing the proposed method a simple experiment was set up in which the original image was processed using the embedding process described in \ref{embedding}.
The results of that process can be seen in figure \ref{fig:lenaEmbedOnceEmbedding}, showing the original image and the image after the embedding process.
Note that in this case the Image Reference content was only embedded once

@@

As an experimental control the results from the experiment in the proposed method will be compared to other current methods for images with self recovering capabilities.
This is discussed further in  section \ref{interpretation}.

\subsection{Interpretation of results}
